This work would not have been possible without the trust and support that Professor Antonio Di Scala has given me for the past two years. Thanks to his enthusiasm, I had the first opportunities to explore the fascinating world of Bitcoin mining, from a pragmatic point of view. Thanks to him, my entire master's degree has been enriched by other academic experiences that have helped me grow tremendously.

I would like to extend a sincere thanks to Alps Blockchain, especially to Francesco, who gave me the opportunity to work on the practical part of this thesis. But more importantly, since the day I learned about their activities, they have introduced me to a world of possibilities where Bitcoin mining can be applied in different real-life contexts. These guys are doing an amazing job, and I am incredibly grateful for the chance to collaborate with them.

A special gratitude goes to Giorgio and Salvatore. Together, we have motivated and inspired each other to create one of the most remarkable experiences that has defined my entire time at Politecnico: BitPolito. Thank you for believing in the mission of our student team to an extent that I could have never imagined before.


Thanks to the entire Trust In Food team. Thank you for being with me during my first-ever startup experience. Without all of you and our shared journey, I wouldn't have developed certain thoughts that help me evaluate every step of my life.

Thanks to all the members of The Pub. Without all of you, I couldn't say that I have lived wonderful years of my life.

Thanks to Push, Jaco, and Gabri for making me rejoice in every day spent in Turin.

Thanks to my entire family, who has always understood me and supported every decision I have made in life.

Thanks to my Irene, who has revolutionized my life and for the past four years has been my driving force to become a constantly improving man.
